---
output:
  pdf_document: 
    fig_width: 6
    fig_height: 4
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Course Practical Assignment - 2nd Deliverable (17 d'abril del 2019)
## *Josep Clotet Ginovart*
## *Eric Martin Obispo*

# Bank client data
## Description of input variables:

  1. age (numeric)
  2. job : type of job (categorical: 'admin', 'blue-collar', 'entrepreneur', 'housemaid', 'management', 'retired', 'self-employed', 'services', 'student', 'technician', 'unemployed', 'unknown')
  3. marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
  4. education (categorical:'basic.4y', 'basic.6y', 'basic.9y', 'high.school', 'illiterate', 'professional.course', 'university.degree', 'unknown')
  5. default: has credit in default? (categorical: 'no','yes','unknown')
  6. housing: has housing loan? (categorical: 'no','yes','unknown')
  7. loan: has personal loan? (categorical: 'no','yes','unknown')# related with the last contact of the current campaign:
  8. contact: contact communication type (categorical:'cellular','telephone')
  9. month: last contact month of year (categorical: 'jan', 'feb', 'mar',..., 'nov', 'dec')
  10. day_of_week: last contact day of the week (categorical:'mon','tue','wed','thu','fri')
  11. duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
  12. campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
  13. pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
  14. previous: number of contacts performed before this campaign and for this client (numeric)
  15. poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')# social and economic context attributes
  16. emp.var.rate: employment variation rate - quarterly indicator (numeric)
  17. cons.price.idx: consumer price index - monthly indicator (numeric)
  18. cons.conf.idx: consumer confidence index - monthly indicator (numeric)
  19. euribor3m: euribor 3 month rate - daily indicator (numeric)
  20. nr.employed: number of employees - quarterly indicator (numeric)
  21. y - has the client subscribed a term deposit? (binary: 'yes','no')

# Loading packages:
```{r, include=FALSE}
# Required Packages: to be increased over the course
requiredPackages <- c("car","knitr","ggplot2","FactoMineR","missMDA","mvoutlier","chemometrics")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]

if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)

```

# Load data from Deliverable 1:
```{r}
load( paste0(dirwd, "/bank-additional/Bank5000_raw.RData") )
summary(df)
```


# PCA:
```{r}
vars_con
res.pca<-PCA(df[, vars_con], quanti.sup=2) #"duration" ha de ser suplementaria ja que es la variable target!
#----> Com que "duration" surt quasi centrada, vol dir que les variables vars_con utilitzades en el PCA no ens ajuden a dir res sobre la variable target.

#----> La variable previous (número de contactes en campanyes antigues) esta relacionada inversament amb pdays (dies que feia que no es trucava el client per altres campanyes).
vars_con2<-c(vars_con[1:3], vars_con[5:10])
res.pca<-PCA(df[, vars_con2], quanti.sup=2) #sense pdays

#veure eixos 3 i 4:
res.pca<-PCA(df[, vars_con2], quanti.sup=2, axes=3:4)

#Dificil extreure conclusions sobre la duració a partir nomes de les variables numèriques!!
summary(res.pca, nb.dec=2, ncp=4)

```
